---
title: "Clustering with DBSCAN"
author: "Suzanne-Kae Rocknathan (1004617)  & Vaishnavi Divya Shridar (1004600)"
date: "Term 7, 2022"
output:
  ioslides_presentation:
    toc: true
    widescreen: true
    smaller: true

---
<style>
.reveal .slides section .slideContent{
    font-size: 10pt;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, message=FALSE, include=FALSE, echo=FALSE, warning=FALSE}
library(factoextra)      # for dataset and visualization
library(ggplot2)         # for visualization
library(stats)           # for kmeans() function (kmeans clustering)
library(cluster)         # for the silhouette() function
library(kernlab)         # for specc() function (spectral clustering)
library(fpc)             # for dbscan()
library(dbscan)          # for dbscan()
library(gifski)   
```


## The Dataset

Given a dataset that looks like this...
```{r, echo=FALSE}
data("multishapes")
df <- multishapes[, 1:2]
#head(df)

ggplot(df,
       aes(x,y)) +
  geom_point()
```

> How would you group these points?

## Let's try clustering it!

Previously, we learned about 2 methods of clustering:  

*   K-means clustering
*   Spectral clustering

Let's take a look at how these methods would group the given dataset.

## K-means clustering plot
```{r, warning = FALSE, echo=FALSE}
set.seed(555)

km <- kmeans(df, 5, nstart = 25)

plot(df, col=km$cluster, main="K-means Clustering")
```
 
The clusters here aren't so accurate. 
Like we learned in class, `kmeans()` usually finds clusters that are spherical and non-overlapping, hence it **doesn't capture nested clusters** like the concentric ovals above.


## Spectral clustering plot

```{r, echo=FALSE}
set.seed(555)
sc <- specc(as.matrix(df), centers = 5, kpar = "automatic")

plot(df, col=sc, main = "Spectral Clustering", frame = FALSE)
```

The spectral clustering output for automatically chosen parameters isn't so accurate either...

## Spectral clustering with sigma = 500

By adjusting the sigma parameters, we can get a better output.
```{r, echo=FALSE}
set.seed(555)
sc <- specc(as.matrix(df), centers = 5, kpar = list(sigma=500))

plot(df, col=sc, main = "Spectral Clustering (sigma = 500)", frame = FALSE)
```

## Results
The results are much better, but not perfect. 

Notice that...   

*   You have to adjust the sigma parameter to obtain the best clusters. 
    *   With multivariate datasets, you won't be able to visualize your plots to check if the clusters look appropriate.
*   The clusters **include noise** (ie. outliers).

*What if we told you there's a better clustering method for this dataset?* 

## DBSCAN clustering plot

Let's take a look at how the DBSCAN method performs!
```{r, warning = FALSE, echo=FALSE}
set.seed(555)
db <- fpc::dbscan(df, eps = 0.15, MinPts = 5)

plot(db, df, main = "DBSCAN", frame = FALSE)
```

This method managed to not only **successfully group** each of the **visible clusters**, but also **identified the 'noise points'** (that is, points that don't belong to any cluster)!


## Clustering with DBSCAN

```{r, animation.hook="gifski", interval=0.1, echo=FALSE, results='hide'}
set.seed(555)
fpc::dbscan(df, eps = 0.15, MinPts = 5, showplot = 1)
```

*Interested to find out how this works? Read our blog to find out more!*    — Team 16


